Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-06 paper_jasa.tex
comments added

The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid. To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data. 
An example lineup from this experiment is shown in Figure \ref{fig:test_continuous}.  This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5. The actual data plot location is 11. \green{Check this} This lineup, we think, is a difficult one. To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters. For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest). For one combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups. Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0. The order that the lineups are shown to participants is randomized.\blue{The order is as described not randomized after. The randomization is done picking first combination of sample size and sd as well as picking lineups for each combination of parameter} \green{Is this a correct description of the procedure?} Each subject saw a substantial number of difficult lineups. The intention was noble, to get a better handle on the power of the lineups for small slopes, but practically this made is very tiring for the participants. It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. \green{perhaps should go into so much detail about this, or keep this for the conclusions section.}

@@

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=n+1, ...,n*c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data. For the experiment we consider $n=100$, $c=0.15$ producing actual data with $115$ points. Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen such that overall error standard deviation of the data is $5$ which is approximately 3.5\green{????}. Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.  A linear model (Equation \eqref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data. This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2. The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null data plots are scatterplots of null data generated by simulating residuals from $N(0, {\hat{\sigma}}^2$) and plotting against $X_1$.

>>>>

The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid. To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data. 
An example lineup from this experiment is shown in Figure \ref{fig:test_continuous}.  This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5. The actual data plot location is 11. \green{Check this} \blue{(Yes, it is 11 !)} This lineup, we think, is a difficult one. To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters. For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest). For one combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups. Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0. The order that the lineups are shown to participants is randomized.\blue{The order is as described not randomized after. The randomization is done picking first combination of sample size and sd as well as picking lineups for each combination of parameter} \green{Is this a correct description of the procedure?} Each subject saw a substantial number of difficult lineups. The intention was noble, to get a better handle on the power of the lineups for small slopes, but practically this made is very tiring for the participants. It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. \green{perhaps should go into so much detail about this, or keep this for the conclusions section.}

@@

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=n+1, ...,n*c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data. For the experiment we consider $n=100$, $c=0.15$ producing actual data with $115$ points. Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen such that overall error standard deviation of the data is $5$ which is approximately 3.5\green{????} \blue{(Yes, this is how we did it. This value (3.5) makes the overall sd to be 5 because of the inclusion of the contaminated part of the data.)}. Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.  A linear model (Equation \eqref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data. This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2. The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null data plots are scatterplots of null data generated by simulating residuals from $N(0, {\hat{\sigma}}^2$) and plotting against $X_1$.
